%\subsection*{Personal}

I had the good fortune to meet and work with a large number of incredibly helpful and talented scientists over the course of my PhD. There are far too many to name here, but I would like to thank everyone I collaborated with, discussed work with, and badgered over email over the last few years. Particular thanks is due to everyone who overcame travel restrictions and a global pandemic to help out during the experiment; it was touch-and-go at a few points, and I'm very glad that it all worked out in the end!

Any PhD student will tell you how dependent the experience is on their supervisor, and I was extremely fortunate to have Peter Norreys as mine. Peter had an unerring belief in me (even when my belief in me had very much erred), and was constantly introducing me to new people and urging me to explore new ideas. He shaped the research project to fit my interests, whether that was the interest in fusion that led to my PhD application, or my desire to do some experimental work. The best thing I can say about him was that even at the hardest times, I knew I could count on him for support. Thank you Peter! I'd also like to mention everyone in the Norreys group past and present (with special thanks to Ben, Marko, and Ramy), along with all those in the wider ALP sub-department. You all made the experience much easier than it otherwise might have been.

I've never really liked long lists of personal acknowledgements, and so I'll keep mine short. Huge thanks is due to all of my family and friends, for the love and support they've given me throughout the years. You all know who you are, and if I need to tell you it here then frankly I'm doing something wrong.


%\subsection*{Institutional}

%If you want to separate out your thanks for funding and institutional support, I don't think there's any rule against it.  Of course, you could also just remove the subsections and do one big traditional acknowledgement section.

%Lorem ipsum dolor sit amet, consectetur adipiscing elit. Ut luctus tempor ex at pretium. Sed varius, mauris at dapibus lobortis, elit purus tempor neque, facilisis sollicitudin felis nunc a urna. Morbi mattis ante non augue blandit pulvinar. Quisque nec euismod mauris. Nulla et tellus eu nibh auctor malesuada quis imperdiet quam. Sed eget tincidunt velit. Cras molestie sem ipsum, at faucibus quam mattis vel. Quisque vel placerat orci, id tempor urna. Vivamus mollis, neque in aliquam consequat, dui sem volutpat lorem, sit amet tempor ipsum felis eget ante. Integer lacinia nulla vitae felis vulputate, at tincidunt ligula maximus. Aenean venenatis dolor ante, euismod ultrices nibh mollis ac. Ut malesuada aliquam urna, ac interdum magna malesuada posuere.